Secure Composition of Robust and Optimising Compilers

### 2023-07-11 popl/main.tex
relwork

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}
% context
To ensure that secure applications do not leak their secrets, they are required to uphold several security properties such as spatial and temporal memory safety as well as criptographic constant time.
% need
Existing work shows how to enforce these properties individually, in an architecture-independent way, by using secure compiler passes that each focus on an individual property.
% task
Unfortunately, given two secure compiler passes that each preserve a possibly-different security property, there is no way to tell what kind of security property will the composition of those secure compilers preserve.

% object
This paper is the first to study what security properties are preserved across the composition of different secure compiler passes.
% findings
Starting from a general theory of property composition for security-relevant properties (such as the aforementioned ones), this paper formalises a secure multi-pass compiler that preserves those security-relevant properties.
% conclusion
Crucially, this paper derives the security of the multi-pass compiler from the composition of the security properties preserved by its individual passes, which include security-preserving as well as optimisation passes.
% 
From an engineering perspective, this is the desirable approach to building secure compilers.

  % \MKin{context}
  % Memory safety necessitates secrecy, since breaking it can lead to an arbitrarily large attack space, i.e., from simply reading a secret up to alteration of control-flow in order to execute arbitrary code.
  % However, secrecy cannot be achieved with just memory safety alone, since, e.g., private data could be leaked by differences in execution time for varying program inputs.
  % This leakage is avoidable by ensuring that programs are cryptographic constant time, i.e., different inputs do not change execution time.
  % Prior work has shown that strategies to ensure cryptographic constant time can be rendered useless, since a compiler may simply optimize them away.
  % \MKin{need}
  % This is why secure compilation is an integral ingredient to achieve security, because a secure compiler guarantees (formally) that properties that hold in the source, also hold in the target. %% mention attacker? the text describes correctness
  % \MKin{task}
  % Unfortunately, it is an open question whether secure compilers can be engineered in a modular way.
  % Ideally, compiler engineers develop two different compilation passes, one that ensures memory safety and another one that is security preserving with respect to cryptographic constant time. %, because this style of engineering is well-established for its simplifying divide-and-conquer approach as well as enhanced reusability.
  % \MKin{object+findings}
  % This paper answers that question positively: The composition of compilers that are secure with respect to some security-properties yields the intersection of these properties.
  % \MKin{object}
  % This article discusses different variants of compositions of secure compilers and instantiates parts of these results in a case-study.
  % To this end, it presents an optimizing, secure compilation chain that preserves both memory safety and cryptographic constant time.
  % \MKin{conclusion}
  % The compositionality results enable modular proofs for the presented, secure compiler. % that is secure with respect to both memory safety and cryptographic constant time.
  % Some results of this work are implemented in the Coq proof assistant.

@@

This section discusses work on robust preservation\MP{
	robust compilation.
	NOT robust preservation
} (\Cref{subsec:relw:seccomprtp}) and on other secure compilation criteria (\Cref{subsec:relw:seccompcrit}).
Since the case study of \Cref{sec:casestud:defs,sec:casestud:rtp} implements measures for preserving \gls{ms} and \gls{cct}, this section then presents relevant related work as well (\Cref{subsec:relw:msmechs,subsec:relw:cctmechs}).

@@

The robust preservation of properties as a compiler-level criterion has been analyzed extensively~\cite{abate2019jour,patrignani2021rsc,abate2021extacc,patrignani2019survey} and thus we build on that framework.
No existing work is concerned with composing robustly safe compilers.
\MP{perhaps say taht we believe this scales to different trace models too and differnet kinds of props, but the ocmposition of monitors becomes tricky}
The work relating robust preservation with universal composability~\cite{patrignani2022universal} is closest to what this paper presents.\MP{why}
% The authors demonstrate a similar compositionality theorem to what is presented here (\Cref{sec:compcomp}) as well as in an earlier version of this work~\cite{kruse2022csc}.
% However, they do not demonstrate the scalability of the approach by means of a case study.

@@

Different mechanisms for \MP{enforcing?}memory safety exist that also consider the secure compilation domain, i.e., have an active attacker model.
For example, the ,,pointers as capabilities'' principle represents pointers as machine-level capabilities~\cite{korashy2021capableptrs}, which behave in a similar fashion to capabilities by means of linear typing~\cite{morrisett2005L3}.
The approach of this paper also uses linear typing, but differs from $L^{3}$~\cite{morrisett2005L3} in the way that functions are not first-class.
Moreover, this paper considers an active attacker, while the work on $L^{3}$ only discusses whole programs and, thus, has no active attacker model.
The instrumentation to ensure memory safety that this paper presents is inspired by seminal work in the literature~\cite{nagarakatte2009soft}.
That work inserts bounds-checks in front of pointer-dereferences and, for this to work, inserts meta-data information on pointer creation.
It\MP{what} also works in a more advanced setting with structured fields accesses and also introduces a table-lookup for pointers that are stored in memory.
This paper only considers pointers to arrays of simple objects\MP{NOT OBJECTS}, i.e., there are no pointers to pointers or structures.
Several other approaches to memory-safety exist in literature, specifically as compiler instrumentations~\cite{akritidis2009baggy,younan2010paricheck,jung2021pico,shankaranarayana2023tailcheck,dhumbumroong2020boundwarden,nam2019framer,zhou2023fatptrs}, hardware-extensions~\cite{kwon2013lowfat,saileshwar2022heapcheck,chen2023flexpointer,kim2023whistle}, or programming language extensions~\cite{elliott2018checkedc,li2022formalcheckedc,jim2002cyclone,elliott2015guilt,west2005cuckoo,weis2019fyr,benoit2019uniqueness}.
What differentiates this work from them is that this work uses known, compiler-based approaches to ensure memory-safety as a means to investigate secure compiler compositions.
This paper does not provide efficient memory-safety, but serves as a theoretical foundation for the secure compilation domain.

@@

The approach to \MP{preserving?enforcing?} cryptographic constant time in this paper is high-level, where a programming language exposes a way to switch the semantics to a data (operand) independent timing mode.
Since identifiers in $\Lscct$ are annotated with a secrecy tag, this approach is similar to others with information flow control.
For example, Vale~\cite{bond2017vale} uses Dafny to ensure constant-time assembly code, while Jasmin~\cite{almeida2017jasmin} makes use of the Coq proof assistant to reject non-constant-time programs.
CT-Wasm~\cite{watt2019ctwasm} enforces constant-timeness by means of a type system.
Different to the approach of this paper, these approaches necessitate that the programmer writes \gls{cct} code.
An approach to allow programmers to write more high-level code is CryptOpt~\cite{kuepper2023cryptopt}, which generates efficient target-code by means of a randomised search.
This paper abstracts over concrete mitigation strategies and simply assumes that there is a flag to switch to a cryptographic-constant time execution mode.
This can be realised by employing the FaCT~\cite{cauligi2019fact} compiler, which translates common non-constant time code patterns to be constant-time, and the data (object) independent timing execution mode of modern processors.

>>>>

%%
%% The abstract is a short summary of the work to be presented in the
%% article.
\begin{abstract}

  \MKin{context}
  Memory safety necessitates secrecy, since breaking it can lead to an arbitrarily large attack space, i.e., from simply reading a secret up to alteration of control-flow in order to execute arbitrary code.
  However, secrecy cannot be achieved with just memory safety alone, since, e.g., private data could be leaked by differences in execution time for varying program inputs.
  This leakage is avoidable by ensuring that programs are cryptographic constant time, i.e., different inputs do not change execution time.
  Prior work has shown that strategies to ensure cryptographic constant time can be rendered useless, since a compiler may simply optimize them away.
  \MKin{need}
  This is why secure compilation is an integral ingredient to achieve security, because a secure compiler guarantees (formally) that properties that hold in the source, also hold in the target. %% mention attacker? the text describes correctness
  \MKin{task}
  Unfortunately, it is an open question whether secure compilers can be engineered in a modular way.
  Ideally, compiler engineers develop two different compilation passes, one that ensures memory safety and another one that is security preserving with respect to cryptographic constant time. %, because this style of engineering is well-established for its simplifying divide-and-conquer approach as well as enhanced reusability.
  \MKin{object+findings}
  This paper answers that question positively: The composition of compilers that are secure with respect to some security-properties yields the intersection of these properties.
  \MKin{object}
  This article discusses different variants of compositions of secure compilers and instantiates parts of these results in a case-study.
  To this end, it presents an optimizing, secure compilation chain that preserves both memory safety and cryptographic constant time.
  \MKin{conclusion}
  The compositionality results enable modular proofs for the presented, secure compiler. % that is secure with respect to both memory safety and cryptographic constant time.
  Some results of this work are implemented in the Coq proof assistant.

@@

This section discusses work on robust compilation (\Cref{subsec:relw:seccomprtp}) and on other secure compilation criteria (\Cref{subsec:relw:seccompcrit}).
Since the case study of \Cref{sec:casestud:defs,sec:casestud:rtp} implements measures for preserving \gls{ms} and \gls{cct}, this section then presents relevant related work as well (\Cref{subsec:relw:msmechs,subsec:relw:cctmechs}).

@@

The robust preservation of properties as a compiler-level criterion has been analyzed extensively~\cite{abate2019jour,patrignani2021rsc,abate2021extacc,patrignani2019survey} and thus we build on that framework.
No existing work is concerned with composing robustly safe compilers.
\MPin{
  perhaps say taht we believe this scales to different trace models too and differnet kinds of props, but the ocmposition of monitors becomes tricky
  \MKin{conclusion?}
}
The work relating robust preservation with universal composability~\cite{patrignani2022universal} is closest to what this paper presents.
The authors demonstrate a similar compositionalty theorem to what is presented here (\Cref{sec:compcomp}) but use it in the context of protocols.
They do not demonstrate the scalability of the approach.
Moreover, they are missing the upper and lower compositions.
% The authors demonstrate a similar compositionality theorem to what is presented here (\Cref{sec:compcomp}) as well as in an earlier version of this work~\cite{kruse2022csc}.
% However, they do not demonstrate the scalability of the approach by means of a case study.

@@

Different mechanisms for enforcing memory safety exist that also consider the secure compilation domain, i.e., have an active attacker model.
For example, the ,,pointers as capabilities'' principle represents pointers as machine-level capabilities~\cite{korashy2021capableptrs}, which behave in a similar fashion to capabilities by means of linear typing~\cite{morrisett2005L3}.
The approach of this paper also uses linear typing, but differs from $L^{3}$~\cite{morrisett2005L3} in the way that functions are not first-class.
Moreover, this paper considers an active attacker, while the work on $L^{3}$ only discusses whole programs and, thus, has no active attacker model.
The instrumentation to ensure memory safety that this paper presents is inspired by Softbounds~\cite{nagarakatte2009soft}.
That work inserts bounds-checks in front of pointer-dereferences and, for this to work, inserts meta-data information on pointer creation.
Softbounds also works in a more advanced setting with structured fields accesses and also introduces a table-lookup for pointers that are stored in memory.
This paper only considers arrays of primitive data, i.e., there are no pointers to pointers or structures.
Several other approaches to memory-safety exist in literature, specifically as compiler instrumentations~\cite{akritidis2009baggy,younan2010paricheck,jung2021pico,shankaranarayana2023tailcheck,dhumbumroong2020boundwarden,nam2019framer,zhou2023fatptrs}, hardware-extensions~\cite{kwon2013lowfat,saileshwar2022heapcheck,chen2023flexpointer,kim2023whistle}, or programming language extensions~\cite{elliott2018checkedc,li2022formalcheckedc,jim2002cyclone,elliott2015guilt,west2005cuckoo,weis2019fyr,benoit2019uniqueness}.
What differentiates this work from them is that this work uses known, compiler-based approaches to ensure memory-safety as a means to investigate secure compiler compositions.
This paper does not provide efficient memory-safety, but serves as a theoretical foundation for the secure compilation domain.

@@

The approach to preserving cryptographic constant time in this paper is high-level, where a programming language exposes a way to switch the semantics to a data (operand) independent timing mode.
Since identifiers in $\Lscct$ are annotated with a secrecy tag, this approach is similar to others with information flow control.
For example, Vale~\cite{bond2017vale} uses Dafny to ensure constant-time assembly code, while Jasmin~\cite{almeida2017jasmin} makes use of the Coq proof assistant to reject non-constant-time programs.
CT-Wasm~\cite{watt2019ctwasm} enforces constant-timeness by means of a type system.
Different to the approach of this paper, these approaches necessitate that the programmer writes \gls{cct} code.
An approach to allow programmers to write more high-level code is CryptOpt~\cite{kuepper2023cryptopt}, which generates efficient target-code by means of a randomised search.
This paper abstracts over concrete mitigation strategies and simply assumes that there is a flag to switch to a cryptographic-constant time execution mode.
This can be realised by employing the FaCT~\cite{cauligi2019fact} compiler, which translates common non-constant time code patterns to be constant-time, and the data (object) independent timing execution mode of modern processors.
